---
output: github_document
title: cesium
---

[![R-CMD-check](https://github.com/goergen95/cesium/workflows/R-CMD-check/badge.svg)](https://github.com/goergen95/cesium/actions)
[![License](https://img.shields.io/badge/License-GPL (>=3)-brightgreen.svg?style=flat)](https://choosealicense.com/licenses/gpl-3.0/)

## Installation

You can install the development version of `cesium` from GitHub with:

```{r installation, eval = FALSE}
# install.packages("remotes")
remotes::install_github("goergen95/cesium")
```

Visit the [usecases](https://github.com/goergen95/cesium/usecases) directory on 
GitHub to explore applications of the `cesium` package.

## Disclaimer

This package is highly experimental, very far from being feature-equal 
compared with other mapping tools available to the R community and very likely
to include breaking-changes in the foreseeable future. In other words, it needs 
your help! If you would like to be able to use [`CesiumJS`](https://cesium.com/platform/cesiumjs/) 
capabilities from within R, please consider contributing to the development of 
this package. You can report bugs, open feature-requests or contribute 
pull-requests to improve the documentation or the code base. We have a 
[Code-of-Conduct](https://github.com/goergen95/cesium/CODE_OF_CONDUCT.md) 
governing the process of contributing to this repository. By participating in 
this project you agree to abide by its terms.

## Getting started

If you are familiar with the [`leaflet`](http://rstudio.github.io/leaflet/) 
package, you will quickly be able to use `cesium` as well. In fact, behind the 
scenes, `cesium` uses the very same design patterns that the authors of leaflet 
did. One distinctive feature of `cesium`, which probably represents the most 
difficult part in using it, is the inclusion of time as another dimension for 
mapping spatial data. It uses [`CesiumJS`](https://cesium.com/platform/cesiumjs/), 
developed by AGI, a Javascript library for mapping spatio-temporal data on an 
interactive 3D globe in the browser. While it also can be used to map data which 
is constant in time, it shows its real strength when the properties of your data 
change over time.

To achieve this, the authors of `CesiumJS` came up with a `JSON` based file format 
called [`CZML`](https://github.com/AnalyticalGraphicsInc/czml-writer/wiki/CZML-Guide). 
The cesium R package, in its essence, is a translator between sf objects to 
`CZML`. This already takes us a long way, because we can use `CZML` to describe 
very different types of geometries in a way that `CesiumJS` can render these 
geometries and their properties dynamically.

To enable you to use the package efficiently, let's briefly discuss some of 
the basic building blocks.

### Basic concepts

The most basic concept in using `cesium` is that of an entity.
An entity is a digital representation of an object located in space and time.
When your data does not change over time, you could think of an entity as a
single row in your `sf` object, where certain measurements, or properties in the
`CesiumJS` lingua, are associated with a specific location in space. This becomes
a little bit more complicated when your data has a temporal component. Now, 
either the entity might change its location in space over time, or it might 
change certain properties over time, or both. To complicate it even further,
if you have a collection of entities that vary over time, more often
than not they are not sampled at the exact same time steps. You end up with a 
time series where potentially both the location and properties of the entities 
vary irregularly over time.


```{r example}
library(sf)
data <- data.frame(
  ID = c(1,1,2,2),
  M = runif(4),
  time = c(Sys.time() - 0:1 * 30,  Sys.time() - 1:2 * 15),
  geom = st_as_sfc(apply(cbind(runif(4, -180, 180), runif(4, -90, 90)), 1, st_point, simplify = F))
)

(data <- st_as_sf(data, crs = st_crs("EPSG:4326")))
```

To accommodate such irregular time series as the one above, we use `sf` as 
the digital representation for such objects. If your data varies over time, 
`cesium` expects you to supply a variable which identifies all observations 
that belong to a given entity. In the case of the data shown above, you would be 
required to set `id_var = "ID"`. Additionally, you need to supply the variable 
that identifies the time of each observation. In the case above, that would
translate to `time_var = "time"`.

As you can see by inspecting the `geometry` column, the two entities change
their location over time. `cesium` does not check if an entity changes its 
location, but instead offers an argument indicating whether or not to treat 
space as constant or dynamic. In case of constant space, the default, the first 
location is used for all subsequent observations of an entity. For the data
above, we thus would have to set `constant_space = FALSE`, explicitly.

In case of time-dynamic entities, the question arises how certain properties 
should be treated *between* the sampled time steps. `cesium` let's you decide 
between two different possibilities:

- intervals: The default behavior in `cesium`. In this case, a property is only 
  valid for an interval between the actual observation and the following one. 
  This means that the value of a property changes abruptly once the respective 
  interval is no longer valid. You do not have to do anything special to 
  enable intervals for time-tagged properties, since this is the default behavior.

- interpolation: In this case, properties are interpolated for the time between 
  observations. **Note**, that interpolation can only be done for numerical 
  properties. It **cannot** be sensually done for things like a text label or 
  an image. You can use `interpolation_options()` in most of the `czml_*()`
  functions to fine-control the interpolation behavior.


### Implementation & Limitations

Currently, `cesium` provides basic support to plot the following items:

- point geometries enabled by `add_points()` and `point_options()`
- markers enabled by `add_markers()` and `marker_options()`
- labels enabled by `add_labels()` and `label_options()`
- line geometries enabled by `add_lines()` and `line_options()`
- polygon geometries enabled by `add_polygons()` and `polygon_options()`
- raster image time series enabled by `add_raster()`

Most arguments should be set using the respective `czml_*()` function. Please
consider the help page if in doubt. If your data has no temporal dimension
to interpolate over or you wish to hold certain properties constant, you might
supply single values that are then applied for all time steps. 

In case you want to vary a property over time, you can use the formula notation 
and additionally supply the column where the time steps are found. **Note**, 
that different from `leaflet`, formulas are resolved on the entity level. If you 
wish to set a property based on a global statistic, you will need to pre-compute 
it.

Following is a list of features that most probably are required for a productive
usage of `cesium` but currently are not implemented:

- conversion methods for `terra` and other spatial classes
- support for `sf` geometries of type `MULTI*`
- mechanism to supply base maps
- layer manager enabling toggling visibility of layers
- legends for colors, sizes, etc.
- support for custom TMS/WMS servers
- support for custom terrain providers
- R Markdown and shiny support

Please contact the maintainers of the package if you would like to support
the development of these or other items.

### Acknowledgments

On the R-side, `cesium` borrows heavily from the `leaflet` package for the 
internal workings of the code and the color mechanisms. 
Without the people driving the development of `CesiumJS` and the `CZML`
format, such a thin wrapper as `cesium` would have no chance to go that far. 

## Licence

This package is licensed to you under the terms of the GNU General Public 
License version 3 or later.

Copyright 2023 Darius A. Görgen
